---
title: "Data Visualization Final Project"
author: "Nathan Colbert"
date: "3/25/2018"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("R Scripts/DVFP Setup.r")
knitr::opts_chunk$set(eval=FALSE)
```
# Rachmaninoff - Complete
## Rachmaninoff 5th Piano Prelude in G minor
### Load in
```{r}
rach_5 <- readMidi("Midi/Rach23n5.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```


### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is Bb (10) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 5
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}
  
## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```



### Piece Information
```{r}
#change info for new pieces
r5_dun <- r5_test2  %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Rachmaninoff",
                              piece = "Prelude Number 5 in G Minor",
                              Tonality = "Minor",
                              Type = "Solo Piano")

```



## Rachmaninoff Second Piano Sonata in Bb Minor
### Load in
```{r}
rach_5 <- readMidi("Midi/Rachson2.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```



### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is Bb (10) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 2
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}
  
## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```



## Change names/variable here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Rachmaninoff",
                              piece = "Piano Sonata 2 in Bb Minor",
                              Tonality = "Minor",
                              Type = "Piano Sonata")

Rachmaninoff <- rbind(r5_dun, r5_dun2)
```


## Rachmaninoff Prelude no. 1 in C major 
### Load in
```{r}
rach_5 <- readMidi("Midi/tmra3201.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```




## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is Bb (10) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```






## Change piece name here
```{r}
#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Rachmaninoff",
                              piece = "Piano Prelude no. 1 in C Major",
                              Tonality = "Major",
                              Type = "Solo Piano")

Rachmaninoff <- rbind(Rachmaninoff, r5_dun2)

```


## Rachmaninoff Prelude no. 11 in B major 
### Load in
```{r}
rach_5 <- readMidi("Midi/tmra3211.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```
## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is B (11) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 1
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```



## Change piece name here
```{r}
#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Rachmaninoff",
                              piece = "Piano Prelude no. 11 in B Major",
                              Tonality = "Major",
                              Type = "Solo Piano")

Rachmaninoff <- rbind(Rachmaninoff, r5_dun2)

```


## Rachmaninoff Romance for 2 Pianos (Cminor) 
### Load in
```{r}
rach_5 <- readMidi("Midi/Rachmani.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```


## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is C (12) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


## Change piece name here
```{r}
#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Rachmaninoff",
                              piece = "Romance for 2 Pianos",
                              Tonality = "Major",
                              Type = "Piano Duet")

Rachmaninoff <- rbind(Rachmaninoff, r5_dun2)
write.csv(Rachmaninoff, "rachmaninoff2.csv")
```


# Tchaikovsky - Complete
## Tchaikovsky Dance of the Sugar Plum Fairy (E minor)
### Load in
```{r}
rach_5 <- readMidi("Midi/fee.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is E (4) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 8
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```



## Change here
```{r}

#change here for new pieces
Tchaikovsky <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                   two_lag_tran = lag(tran_chord_named, n = 2),
                   three_lag_tran = lag(tran_chord_named, n = 3),
                   four_lag_tran = lag(tran_chord_named, n = 4),
                   composer = "Tchaikovsky",
                   piece = "Dance of the Sugar Plum Fairy",
                   Tonality = "Minor",
                   Type = "Ballet")

```


## Tchaikovsky String Quartet No.1 in D, Op.11 (D Major)
### Load in
```{r}
rach_5 <- readMidi("Midi/tch_qr1a.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```

## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is D (2) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 10
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```



## Change names/variable here
```{r}
#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Tchaikovsky",
                              piece = "String Quartet no. 1 in D, Mvt 1",
                              Tonality = "Major",
                              Type = "String Quartet")

Tchaikovsky <- rbind(Tchaikovsky, r5_dun2)
```


## Tchaikovsky Swan Lake (Valse) (A major)
### Load in
```{r}
rach_5 <- readMidi("Midi/swanlk-2.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')
```

## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is A (8) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 4
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')
```


## Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Tchaikovsky",
                              piece = "Swan Lake (Valse)",
                              Tonality = "Major",
                              Type = "Ballet")

Tchaikovsky <- rbind(Tchaikovsky, r5_dun2)

```


## Tchaikovsky Piano Concerto 1 movement 1 (Bb minor)
### Load in
```{r}
rach_5 <- readMidi("Midi/tchai1-1.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is Bb (10) and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 2
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


## Change piece name here
```{r}
#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Tchaikovsky",
                              piece = "Piano Concerto 1, mvt. 1",
                              Tonality = "Minor",
                              Type = "Piano Concerto")

Tchaikovsky <- rbind(Tchaikovsky, r5_dun2)

```


## Tchaikovsky Romance (F Minor)
### Load in
```{r}
rach_5 <- readMidi("Midi/Romancef.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

## Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 7
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


## Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Tchaikovsky",
                              piece = "Romance in F minor",
                              Tonality = "Minor",
                              Type = "Solo Piano")

Tchaikovsky <- rbind(Tchaikovsky, r5_dun2)
write.csv(Tchaikovsky, "tchaikovsky2.csv")
```

# Chopin - Complete
## Ballade Number 1 in G Minor
### Load in
```{r}
rach_5 <- readMidi("Midi/chop23b.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 7  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 5
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
Chopin <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Chopin",
                              piece = "Ballade No 1 in G Minor",
                              Tonality = "Minor",
                              Type = "Solo Piano")


```

## Nocturne No.19 in E-
### Load in 
```{r}
rach_5 <- readMidi("Midi/op72.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 4 and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 8
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Chopin",
                              piece = "Nocturne No.19 in E-",
                              Tonality = "Minor",
                              Type = "Solo Piano")

Chopin <- rbind(Chopin, r5_dun2)
```

## Scherzo No.2 in Bb-
### Load in
```{r}
rach_5 <- readMidi("Midi/choschz2.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 10  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 2
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Chopin",
                              piece = "Scherzo No.2 in Bb-",
                              Tonality = "Minor",
                              Type = "Solo Piano")

Chopin <- rbind(Chopin, r5_dun2)
```

## Grand Duo on Themes from Robert Le Diable
### Load in
```{r}
rach_5 <- readMidi("Midi/gdcon.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 1 and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 1
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Chopin",
                              piece = "Grand Duo on Themes from Robert Le Diable",
                              Tonality = "Minor",
                              Type = "Duet")

Chopin <- rbind(Chopin, r5_dun2)
```

## Ballade No.4 in F-
### Load in
```{r}
rach_5 <- readMidi("Midi/chopin-1.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 6  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 7
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Chopin",
                              piece = "Ballade No.4 in F-",
                              Tonality = "Minor",
                              Type = "Solo Piano")

Chopin <- rbind(Chopin, r5_dun2)
View(Chopin)

write.csv(Chopin, "chopin.csv")
```


# Brahms - Complete
## String Quintet No.2 in G, Mvt. 2 Adagio
### Load in
```{r}
rach_5 <- readMidi("Midi/jbqu2ii.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 4  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 8
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
Brahms <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Brahms",
                              piece = "String Quintet No.2 in G, Mvt. 2 Adagio",
                              Tonality = "Minor",
                              Type = "String Quintet")


```

## Ein deutsches Requiem, 4.Chorus: Wie lieblich sind deine Wohnungen (Eb Major)
### Load in
```{r}
rach_5 <- readMidi("Midi/brmsreq4.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 9
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Brahms",
                              piece = "Ein deutsches Requiem, 4.Chorus: Wie lieblich sind deine Wohnungen",
                              Tonality = "Major",
                              Type = "Choral")

Brahms <- rbind(Brahms, r5_dun2)
```

## Symphony No.3 in F, Mvt. 2 Andante
### Load in
```{r}
rach_5 <- readMidi("Midi/jk_br90b.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Brahms",
                              piece = "Symphony No.3 in F, Mvt 2 Andante",
                              Tonality = "Major",
                              Type = "Symphony")

Brahms <- rbind(Brahms, r5_dun2)
```

## Piano Sonata No.1 in C, Mvt. 1 Allegro
### Load in
```{r}
rach_5 <- readMidi("Midi/jbop1n1.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Brahms",
                              piece = "Piano Sonata No.1 in C, Mvt. 1 Allegro",
                              Tonality = "Major",
                              Type = "Solo Piano")

Brahms <- rbind(Brahms, r5_dun2)
```

## 11 Chorale Preludes, 1.Mein Jesu, der du mich
### Load in
```{r}
rach_5 <- readMidi("Midi/jbmj1221.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 8
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Brahms",
                              piece = "11 Chorale Preludes, Op.posth.122	 
1.Mein Jesu, der du mich",
                              Tonality = "Minor",
                              Type = "Organ")
Brahms <- rbind(Brahms, r5_dun2)
View(Brahms)
write.csv(Brahms, "brahms.csv")
```

# Berlioz - In Progress

## Symphonie Fantastique: 4. Marche au supplice (G minor)
### Load in
```{r}
rach_5 <- readMidi("Midi/syfan4.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 5
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
Berlioz <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Berlioz",
                              piece = "Symphonie Fantastique: 4. Marche au supplice",
                              Tonality = "Minor",
                              Type = "Symphony")



write.csv(Berlioz, "berlioz.csv")
```


# Beethoven - Complete
## Symphony No.5 in C-, Op.67	, 1.Allegro con brio
### Load in
```{r}
rach_5 <- readMidi("Midi/dglvb67a.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Beethoven",
                              piece = "Symphony 5, mvt. 1",
                              Tonality = "Minor",
                              Type = "Symphony")

```

## String Quartet No.13 in Bb, Op.130 ('Lieb'), 1.Adagio ma non troppo. Allegro
### Load in
```{r}
rach_5 <- readMidi("Midi/beesq131.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 10  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i]  + 2
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun3 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Beethoven",
                              piece = "String Quartet 13, mvt. 1",
                              Tonality = "Major",
                              Type = "String Quartet")

Beethoven <- rbind(r5_dun2, r5_dun3)
```

## Symphony No.7 in A, Op.92, 4.Finale: Allegro con brio
### Load in
```{r}
rach_5 <- readMidi("Midi/b7-4.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 9  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 3
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Beethoven",
                              piece = "Symphony 7, mvt. 4",
                              Tonality = "Major",
                              Type = "Symphony")


Beethoven <- rbind(Beethoven, r5_dun2)
```

## Piano Sonata No.10 in G, Op.14, No.2, 2.Andante
### Load in
```{r}
rach_5 <- readMidi("Midi/bps10-2.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 12  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Beethoven",
                              piece = "Piano Sonata 10 mvt. 2",
                              Tonality = "Major",
                              Type = "Solo Piano")



Beethoven <- rbind(Beethoven, r5_dun2)
```

## Piano Concerto No.1 in C, Op.15, 
### Load in
```{r}
rach_5 <- readMidi("Midi/be-pc-01.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Beethoven",
                              piece = " Piano Concerto 1",
                              Tonality = "Major",
                              Type = "Piano Concerto")


Beethoven <- rbind(Beethoven, r5_dun2)

write.csv(Beethoven, "beethoven.csv")
```

# Verdi - Complete
## Ave Maria, for mezzo-soprano, chorus, and orchestra	 
### Load in
```{r}
rach_5 <- readMidi("Midi/ver_avem.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 12  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Verdi",
                              piece = "Ave Maria",
                              Tonality = "Major",
                              Type = "Symphony")


```

## Alzira (tragedia lirica), 1.Sinfonia
### Load in
```{r}
rach_5 <- readMidi("Midi/ajsalz00.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 2  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] +10
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Verdi",
                              piece = "Alzira (tragedia lirica), 1.Sinfonia",
                              Tonality = "Minor",
                              Type = "Opera")

Verdi <- rbind(r5_dun, r5_dun2)
```

## Aida (opera), 1.Prelude
### Load in
```{r}
rach_5 <- readMidi("Midi/verdi_a_.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 2  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 10
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Verdi",
                              piece = "Aida, 1.Prelude",
                              Tonality = "Major",
                              Type = "Opera")


Verdi <- rbind(Verdi, r5_dun2)
```

## La forza del destino, 1.Sinfonia (Overture)
### Load in
```{r}
rach_5 <- readMidi("Midi/vfdesou.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 9  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 3 
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}


#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer ="Verdi",
                              piece = "La forza del destino (opera), 1.Sinfonia",
                              Tonality = "Minor",
                              Type = "Opera")

Verdi <- rbind(Verdi, r5_dun2)
```

## Requiem Mass, for soloists, chorus, and orchestra (Manzoni Requiem), 1.Dies irae
### Load in
```{r}
rach_5 <- readMidi("Midi/ingemi1.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 7  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 5
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Verdi",
                              piece = "Requiem, Dies Irae",
                              Tonality = "Minor",
                              Type = "Symphony")

Verdi <- rbind(Verdi, r5_dun2)
write.csv(Verdi, "verdi.csv")
```

# Dvorak - Only 3 pieces
## Symphony no.9 Mvt. 2 Largo
### Load in
```{r}
rach_5 <- readMidi("Midi/dvolargo.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 1  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 11
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Dvorak",
                              piece = "Symphony 9, Mvt. 2 Largo",
                              Tonality = "Major",
                              Type = "Symphony")



```

## Piece name here
### Load in
```{r}
rach_5 <- readMidi("Midi/polrus.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 0
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun3 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Dvorak",
                              piece = "Rusalka, 5.Ballet: Slavnostní hudba (Polonaise)",
                              Tonality = "Minor",
                              Type = "Ballet")

Dvorak <- rbind(r5_dun2, r5_dun3)
```

## String Quartet No.12 in F ('American'), Op.96, 1.Allegro ma non troppo
### Load in
```{r}
rach_5 <- readMidi("Midi/dv96_1.mid")

## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 5  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 7
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Dvorak",
                              piece = "String Quartet 12, mvt. 1",
                              Tonality = "Major",
                              Type = "String Quartet")


Dvorak <- rbind(Dvorak, r5_dun2)

write.csv(Dvorak, "dvorak.csv")
```


# Mahler - Only 3 Pieces
## Symphony No.5 in C#-, 5.Rondo-Finale: Allegro
### Load in
```{r}
rach_5 <- readMidi("Midi/mahler55.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 2  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 10
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Mahler",
                              piece = "Symphony No.5 in C#-, 5.Rondo-Finale: Allegro",
                              Tonality = "Major",
                              Type = "Symphony")


```

## Symphony No.9 in D, 1.Andante comodo
### Load in
```{r}
rach_5 <- readMidi("Midi/ma_9_1.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 2  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 10
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Mahler",
                              piece = "Symphony No.9 in D, 1.Andante comodo",
                              Tonality = "Major",
                              Type = "Symphony")

Mahler <- rbind(r5_dun, r5_dun2)
```

## Symphony No.6 in A- ('Tragic'), 1.Allegro energico, ma non troppo
### Load in
```{r}
rach_5 <- readMidi("Midi/mahler61.mid")
## Midi Transformations/Octave Normalizing
source('R Scripts/MidiTransform.r')

```

### Change Key Here
```{r}
# take key num and add appropriate amount to transpose to C, here key is 9  and c is 12
for (i in 1:NROW(r5$note)) {
  r5$tran_note[i] = r5$note[i] + 3
  if (r5$tran_note[i] > 12){
    r5$tran_note[i] = r5$tran_note[i] - 12
  } else {
    r5$tran_note[i] = r5$tran_note[i]
  }
}

## Chord and Roman Numeral Loops
source('R Scripts/C_Loop.r')

```


### Change piece name here
```{r}

#change here for new pieces
r5_dun2 <- r5_test2 %>% mutate(one_lag_tran = lag(tran_chord_named, n = 1),
                              two_lag_tran = lag(tran_chord_named, n = 2),
                              three_lag_tran = lag(tran_chord_named, n = 3),
                              four_lag_tran = lag(tran_chord_named, n = 4),
                              composer = "Mahler",
                              piece = "Symphony No.6, Mvt. 1",
                              Tonality = "Minor",
                              Type = "Symphony")

Mahler <- rbind(Mahler, r5_dun2)
write.csv(Mahler, "mahler.csv")
```

