---
title: "Data Visualization Assignment 1"
author: "Nathan Colbert"
date: "2/14/2018"
output:
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(reshape2)
library(ggthemes)
library(grid)
library(png)
library(rvest)
library(stringr)
library(ggplot2)
library(plotly)
library(grid)
library(ggrepel)
library(DT)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}

winter <- read.csv("Desktop/winter.csv")
dictionary <- read.csv("Desktop/dictionary.csv")
Winter <- rename(winter, Code = Country)
Winter$Code[Winter$Code == "GDR"] <- "GER"
Winter$Code[Winter$Code == "EAU"] <- "GER"
Winter$Code[Winter$Code == "FRG"] <- "GER" #How could I have done this better?
Winter$Code[Winter$Code == "URS"] <- "RUS"
Winter$Code[Winter$Code == "EUN"] <- "RUS"
Winter$Code[Winter$Code == "RU1"] <- "RUS"

olympics <- full_join(Winter, dictionary)
```

# Medal Count Over Time

```{r, echo=FALSE, message=FALSE, fig.width=10}

gold_medal <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Gold") %>% group_by(Country) %>% count(Medal)

silver_medal <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Silver") %>% group_by(Country) %>% count(Medal)

bronze_medal <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Bronze") %>% group_by(Country) %>% count(Medal)



by_country_medal <- full_join(gold_medal, silver_medal, by = "Country")

by_country_medal <- full_join(by_country_medal, bronze_medal, by = "Country")


by_country_medal <- by_country_medal %>% rename("Bronze" = n) %>% rename("Silver" = n.y) %>% rename("Gold" = n.x)  %>% replace_na(list(Gold = 0, Silver = 0, Bronze = 0)) %>% arrange(desc(Gold)) %>% head(n = 10)
test <- by_country_medal %>%  melt() %>% select(Country, variable, value) %>% arrange(Country)


ggplot(test, aes(x = Country, y = value, fill = variable, name = "Medal Type")) + geom_bar(stat = "identity", position = "dodge") + ylab("Medals Won") + xlab(NULL) + scale_fill_manual("variable", values = c("Gold"= "tan", "Silver" = "light grey", "Bronze" = "#A57164")) + theme_tufte() + coord_flip() + theme(legend.title = element_blank(), plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) +labs(title = "Top Ten Most Successful Countries in the Winter Olympics") 


gold_medal_2 <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Gold") %>% group_by(Country, Year) %>% count(Medal)

silver_medal_2 <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Silver") %>% group_by(Country, Year) %>% count(Medal)

bronze_medal_2 <- olympics %>% select(Country, Medal, Year) %>% filter(Medal == "Bronze") %>% group_by(Country, Year) %>% count(Medal)

by_country_year <- full_join(gold_medal_2, silver_medal_2, by = c("Country", "Year"))

by_country_year <- full_join(by_country_year, bronze_medal_2, by = c("Country", "Year"))
by_country_year <- by_country_year %>% rename("Bronze" = n) %>% rename("Silver" = n.y) %>% rename("Gold" = n.x)  %>% replace_na(list(Gold = 0, Silver = 0, Bronze = 0, Medal.x = "Gold", Medal.y = "Silver", Medal = "Bronze")) %>% arrange(desc(Gold))


top_ten <- c("United States", "Switzerland", "Sweden", "Russia", "Norway", "Italy", "Germany", "Finland", "Canada", "Austria")
test2 <- by_country_year %>% melt(id.vars = c("Year", "Country"), measure.vars = c("Gold", "Silver", "Bronze")) %>% filter(Country %in% top_ten) %>% arrange(Year)


medal_time <- ggplot(test2, aes(x = Year, y = value)) + geom_point(aes(group = Country)) + ylab("Total Medals Won") + xlab(NULL) + theme_tufte() + geom_line(aes(color = Country))+ facet_wrap(~ Country, nrow = 2) + theme(legend.title = element_blank(), plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) + labs(title = "Medals Won by Top Ten Countries Over Time")
medal_time
```

Of these two graphs, I would recommend my editor use the bar chart. Even though the line chart is more visually compelling, the bar chart is more easily discerned. 


# Medal Counts Adjusted by Population and GDP

```{r, echo=FALSE, message=FALSE, fig.width=10}
by_pop <- olympics %>% select(Country, Medal, Population, GDP.per.Capita) %>% filter(Medal == "Gold") %>% group_by(Country, Population, GDP.per.Capita) %>% count(Medal) %>% mutate(Pop_Scale = n / Population, GDP_scale = n / GDP.per.Capita) %>% arrange(desc(Pop_Scale)) %>% head(n = 10)

by_GDP <- olympics %>% select(Country, Medal, Population, GDP.per.Capita) %>% filter(Medal == "Gold") %>% group_by(Country, Population, GDP.per.Capita) %>% count(Medal) %>% mutate(Pop_Scale = n / Population, GDP_scale = n / GDP.per.Capita) %>% arrange((GDP_scale)) %>% head(n = 10)

ggplot(by_pop, aes(x = reorder(Country, Pop_Scale), y = Pop_Scale * 100000)) + geom_bar(stat = "identity", fill = "tan") + coord_flip() + xlab(NULL) + ylab("Gold Medal Count Scaled by Population") + theme_tufte() + theme(legend.title = element_blank(), plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) + labs(title = "Top Ten Most Successful Countries When Scaled by Population")
  
ggplot(by_GDP, aes(x = reorder(Country, GDP_scale), y = GDP_scale * 10000)) + geom_bar(stat = "identity", fill = "tan") + coord_flip() + xlab(NULL) + ylab("Gold Medal Count Scaled by GDP") + theme_tufte() + theme(legend.title = element_blank(), plot.title = element_text(size = 18, face = "bold", hjust = 0.5)) + labs(title = "Top Ten Most Successful Countries When Scaled by GDP")

```

# Host Country Advantage?

```{r, fig.width=10, echo=FALSE, message=FALSE}

wiki_hosts <- read_html("https://en.wikipedia.org/wiki/Winter_Olympic_Games")
hosts <- html_table(html_nodes(wiki_hosts, "table")[[5]], fill=TRUE)
hosts <- hosts[-1,1:3]
hosts$city <- str_split_fixed(hosts$Host, n=2, ",")[,1]
hosts$country <- str_split_fixed(hosts$Host, n=2, ",")[,2]

home <- c("France", "Switzerland", "United States", "Germany", "Norway", "Italy", "Austria", "Japan", "Yugoslavia", "Canada", "Russia")

hosts$Year <- as.integer(hosts$Year)
olympics2 <- full_join(olympics, hosts)

olympics2 <- olympics2 %>% rename("Host City" = city) %>% rename("Host Country" = country)
char_country <- olympics2 %>% filter(Country %in% home) %>% filter(Year != c(1940, 1944) & Year != c(2018, 2022)) %>% select(Country)
char_country<-as.character(unlist(char_country))

char_country2 <- as.character(unlist(as.data.frame(char_country)))


hc_char <- olympics2%>% filter(Country %in% home) %>% filter(Year != c(1940, 1944) & Year != c(2018, 2022)) %>% select('Host Country') %>% unlist()

attr(hc_char, "names") <- NULL
hc_char[hc_char == " France"] <- "France"
hc_char[hc_char == " United States"] <- "United States"
hc_char[hc_char == " Switzerland"] <- "Switzerland"
hc_char[hc_char == " Japan"] <- "Japan" #Is there a better way to remove this space?
hc_char[hc_char == " Norway"] <- "Norway"
hc_char[hc_char == " Italy"] <- "Italy"
hc_char[hc_char == " Austria"] <- "Austria"
hc_char[hc_char == " Germany"] <- "Germany"
hc_char[hc_char == " Russia"] <- "Russia"
hc_char[hc_char == " Canada"] <- "Canada"

host_adv <- olympics2 %>% filter(Country %in% home) %>% filter(Year != c(1940, 1944) & Year != c(2018, 2022)) %>% select(Year, Gender, Medal, Country, Games, `Host Country`) %>% mutate(at_home = hc_char == char_country)

host_plot <-ggplot(host_adv, aes(x = Games, y = Medal, fill = Medal, alpha = at_home)) + geom_bar(stat = "identity") + scale_fill_manual("Medal", values = c("Gold"= "tan", "Silver" = "light grey", "Bronze" = "#A57164")) + scale_alpha_discrete(range = c(.2,.8)) + facet_wrap(~Country, nrow = 5, scales = "free_y") +theme_tufte() +theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.position = "none", plot.title = element_text(size = 18, face = "bold", hjust = .5)) + ylab(NULL) + labs(title = "Medal Count Highlighted When Competing and Hosting")


host_plot

```


# Country Success by Sport/Discipline/Event

```{r, echo=FALSE, message=FALSE, fig.width=10}

us_ski <- olympics %>% filter(Country == 'United States' & Sport == 'Skiing') 
ggplot(us_ski, aes(x = Event, y = Medal, fill = Medal)) + geom_bar(stat = "identity") + facet_grid( ~ Gender) + scale_fill_manual("Medal", values = c("Gold"= "tan", "Silver" = "light grey", "Bronze" = "#A57164")) + coord_flip() + theme_tufte() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "none", plot.title = element_text(size = 18, hjust = .5, face = "bold")) + xlab(NULL) + ylab(NULL) +labs(title = "United States Sucess in Skiing")
```

# Most Successful Atheletes of All Time

```{r, echo=FALSE, message=FALSE, fig.width=10}
athletes <- olympics %>% select(Athlete, Medal, Country, Sport, Gender) %>% group_by(Athlete, Country, Sport, Gender) %>% filter(Medal == "Gold" ) %>% count(Medal) %>%  arrange(desc(n)) %>% head(n = 25)


ggplot(athletes, aes(x = reorder(Athlete, n), y = n, color = Country)) + geom_point(shape = 21, size = 5, stroke = 1)  + theme_tufte() + ylab("Gold Medals Won") + xlab(NULL) + coord_flip()  + labs(title = "Most Successful Atheletes of All Time") + theme( plot.title = element_text(size = 18, hjust = .5, face = "bold"))  +scale_shape(solid = F)

```

# Making Plots interactive


```{r, fig.width=10, message=FALSE, echo=FALSE}
ggplotly(medal_time)

ggplotly(host_plot)
```

Making these two graphs significantly improves the amount and accuracy of information these graphs can deliver. 

# Data Table

```{r, fig.width=10, echo=FALSE}
options(DT.options = list(pageLength = 5, scrollX = TRUE, scrollY = "300px"))
datatable(olympics)

```
I chose to make a data table that contains all the information used in this study.

